### 2025-04-29 Extra/undergraduate-research-symposium-images.qmd
stash

```{r}
 mplot <- function(x, ..., fill_colour = "black", title = NULL, row_labels = NULL, col_labels = NULL, palette_colors = NULL) {
   d <- data.frame(
     x = rep(1:ncol(x), each = nrow(x)),
     y = rep(nrow(x):1, ncol(x)),
     z = as.integer(x)
   )

   if (!is.null(palette_colors)) {
     d$palette_colors <- rep(palette_colors, ncol(x))
   }

   if (is.null(col_labels)) col_labels <- colnames(x)
   if (is.null(row_labels)) row_labels <- rev(rownames(x))

   p <- ggplot(d, aes(x, y)) +
     geom_tile(col = "grey50", 
               linewidth = 0.5, 
               aes(fill = ifelse(z == 1, palette_colors, "white"))) +
     scale_fill_identity() +  
     coord_equal(expand = FALSE) +
     scale_x_continuous(
       breaks = 1:ncol(x),
       labels = col_labels,
       position = "top"
     ) +
     scale_y_continuous(
       breaks = 1:nrow(x),
       labels = row_labels
     ) +
     theme(
       axis.text = element_text(size = rel(0.3)),
       axis.ticks = element_blank(),
       legend.position = "none",
       plot.title = element_text(hjust = 1)
     ) +
     labs(x = "", y = "") +
     rotate_x_text(angle = 90, hjust = 0, vjust = 0.5)

   p
 }

 increase_focus <- function(x, group = NULL) {
   mcp <- apply(
     x, 
     MARGIN = 1, 
     FUN = \(z) { mean(which(z == 1)) }, 
     simplify = FALSE
   ) |> unlist()
   row_order <- order(mcp, na.last = TRUE)
   x <- x[row_order, ]

   if (!is.null(group)) {
     group <- group[row_order]
   }

   mrp <- apply(
     x, 
     MARGIN = 2, 
     FUN = \(z) { mean(which(z == 1)) }, 
     simplify = FALSE
   ) |> unlist()
   x <- x[, order(mrp, na.last = TRUE)]

   list(matrix = x, group = group)
 }

 concentrate <- function(x, group = NULL, max_iter = 100) {
   old <- x
   not_identical <- TRUE
   iter <- 0

   while (not_identical && iter < max_iter) {
     result <- increase_focus(old, group)
     new <- result$matrix
     group <- result$group
     not_identical <- !identical(old, new)
     old <- new
     iter <- iter + 1
   }

   if (iter == max_iter) {
     warning("Reached maximum iterations without convergence.")
   }

   list(matrix = new, group = group)
 }

 produce_clusters <- function(artifact_data, artifact_data_unique, method = "seriation", color_palette = c("#8DD3C7", "#FFFFB3")){

   artifact_data <- artifact_data %>% 
     select_if(~ !is.numeric(.) || sum(.) != 0) %>% 
     mutate_all(~ as.numeric(. > 0))

   jac <- vegdist(artifact_data, "jaccard", binary = TRUE)
   dm <- as.matrix(jac)

   if(method == "seriation"){
   artifact_matrix <- as.matrix(artifact_data)
   rownames(artifact_matrix) <- rownames(artifact_data)

   group <- artifact_data_unique$group
   unique_group <- unique(group)
  # color_palette <-  RColorBrewer::brewer.pal(length(unique_group), "Set3")
   row_colors <- setNames(color_palette, unique_group)
   result <- concentrate(artifact_matrix, group = group)
   concentrated_matrix <- result$matrix
   reordered_group <- result$group
   group_colors <- row_colors[reordered_group]

   final_plot <- mplot(concentrated_matrix,
     palette_colors = group_colors)
   }

   if(method == "network"){
   disim <- 1 - dm
   disim[disim<0.2] <- 0
   disim_net <- network(disim,
                        directed=F,
                        ignore.eval=F,
                        names.eval='weight')
   disim_net %v% 'vertex.names' <- row.names(artifact_data)
   set.seed(500)

   final_plot <- ggraph(disim_net, layout="fr") +
       geom_edge_link(aes(width = weight, 
                          alpha=weight), 
                      edge_colour = "black", 
                      show.legend= F,
                     ) +
       scale_edge_width(range=c(0.5,1.5)) +
       scale_edge_colour_gradient(low = "#CCCCCC",
                                  high = "#000000") +
       geom_node_point(alpha=1, size=5) +
       geom_node_text(aes(label=row.names(artifact_data)), 
                      size=3, 
                      family= "Arial", 
                      repel=T, max.overlaps = 100) +
       theme(text= element_text(size = 15), 
             panel.background = element_rect(fill = 'white')) +
     border(color = "black", size = 0.5)

   }
   return(final_plot)
 }

```

```{r}
seriation_trans <- produce_clusters(trans_artifact_data, 
                                    trans_unique_data, 
                                    method = "seriation")

seriation_proto <- produce_clusters(proto_artifact_data,  
                                    proto_unique_data, 
                                    color_palette = c("#FFFFB3", "#8DD3C7"),
                                    method ="seriation")

seriation_early <- produce_clusters(early_artifact_data, 
                                    early_unique_data, 
                                    method = "seriation")

seriation_evolved <- produce_clusters(evolved_artifact_data, 
                                      evolved_unique_data, 
                                      method = "seriation")
```

```{r}
network_trans <- produce_clusters(trans_artifact_data, trans_unique_data, method = "network")
network_proto <- produce_clusters(proto_artifact_data, proto_unique_data, method = "network")
network_early <- produce_clusters(early_artifact_data, early_unique_data, method = "network")
network_evolved <- produce_clusters(evolved_artifact_data, evolved_unique_data, method = "network")

```

```{r}
ser_net_trans <- plot_grid(seriation_trans, network_trans, align = "hv", labels = "AUTO", rel_widths = c(1.3, 1.4))
ser_net_proto <- plot_grid(seriation_proto, network_proto, align = "hv", rel_widths = c(1.4, 1.4))
ser_net_early <- plot_grid(seriation_early, network_early, align = "hv", rel_widths = c(1.65, 1.8))
ser_net_evo <- plot_grid(seriation_evolved, network_evolved, align = "hv", rel_widths = c(1.4, 1.4))

final <- plot_grid(ser_net_trans, ser_net_proto, ser_net_early, ser_net_evo, align = "hv", ncol = 1, labels = c("Transitional", "Proto", "Early", "Evolved"))

ggsave(final, 
      filename =  "poster.png",
       bg = "white", h = 9, w = 4.5)
```

```{r}
ser_net_trans <- plot_grid(seriation_trans, network_trans, align = "hv", labels = "AUTO", ncol = 1)
ser_net_proto <- plot_grid(seriation_proto, network_proto, align = "hv", ncol = 1)
ser_net_early <- plot_grid(seriation_early, network_early, align = "hv", ncol = 1)
ser_net_evo <- plot_grid(seriation_evolved, network_evolved, align = "hv", ncol = 1)

final <- plot_grid(ser_net_trans, ser_net_proto, ser_net_early, ser_net_evo, align = "hv", ncol = 4, labels = c("Transitional", "Proto", "Early", "Evolved"))

ggsave(final, 
      filename =  "poster2.png",
       bg = "white", w = 12)

```

```{r}
ser_plot <- plot_grid(seriation_trans, seriation_proto, seriation_early, seriation_early, align = "hv", ncol = 1, labels = c("Transitional", "Proto", "Early", "Evolved"), label_size = 9)
net_plot <- plot_grid(network_trans, network_proto, network_early, network_evolved, align = "hv", ncol = 1)

  final <- plot_grid(ser_plot, net_plot, align = "hv", ncol = 2, rel_widths = c(1.3, 1), labels = "AUTO", label_x = c(0, -0.07))

ggsave(final, 
      filename =  "poster3.png", bg = "white", w = 5, h = 7)
```